---
format:
  docx:
    fig-dpi: 300
    fig-width: 7.5
    reference-doc: scoe-are-report.docx
editor: visual
params:
  lea: Fairfield-Suisun Unified
---

```{r setup}
#| include: false

library(tidyverse)
library(flextable)
library(ggrepel)
library(ggtext)
library(here)

if (interactive()) {
  params <- rmarkdown::yaml_front_matter(
    rstudioapi::getSourceEditorContext()$path
  )$params
}

dashboard <- read_csv(here("data/solano_dashboard.csv")) |>
  filter(
    rtype == "D" &
      districtname == params$lea |
      charter_flag == "Y" & rtype == "S" & schoolname == params$lea
  ) |>
  select(
    reportingyear,
    districtname,
    schoolname,
    student_group_long,
    indicator,
    currdenom,
    currstatus,
    color,
    change,
    indicator_eligible
  )

indicator_colors = c(
  "ELA" = "#CC6677",
  "Math" = "#882255",
  "ELPI" = "#AA4499",
  "college/career" = "#332288",
  "graduation" = "#88CCEE",
  "absenteeism" = "#44AA99",
  "suspension" = "#117733",
  "science" = "#DDCC77"
)

da_groups <-
  dashboard |>
  mutate(reportingyear = as_factor(reportingyear)) |>
  filter(
    reportingyear %in% c(2025, 2024, 2023, 2022, 2019, 2018, 2017),
    indicator_eligible == TRUE
  ) |>
  select(-indicator_eligible)

wide_da_groups <-
  da_groups |>
  select(
    reportingyear,
    districtname,
    schoolname,
    student_group_long,
    indicator
  ) |>
  arrange(reportingyear) |>
  pivot_wider(
    names_from = reportingyear,
    values_from = indicator,
    values_fn = list
  ) |>
  arrange(districtname, schoolname, student_group_long)

clean_groups <-
  wide_da_groups |>
  mutate(
    across(where(is.list), ~ map_chr(., ~ paste(., collapse = "; "))),
    da_years = rowSums(across(matches("^\\d{4}$"), ~ . != ""))
  ) |>
  filter(da_years >= 3) |>
  mutate(`2019` = if ("2019" %in% names(wide_da_groups)) `2019` else "") |>
  select(districtname, student_group_long, `2022`, `2023`, `2024`, `2025`)

```

```{r create_variables}
#| include: false

student_groups_list <- unique(clean_groups$student_group_long)
student_group_count <- length(student_groups_list)

initials <-
  switch(
    params$lea,
    "Solano County Office of Education" = "SCOE",
    "Benicia Unified" = "BUSD",
    "Dixon Unified" = "DUSD",
    "Fairfield-Suisun Unified" = "FSUSD",
    "Travis Unified" = "TUSD",
    "Vacaville Unified" = "VUSD",
    "Vallejo City Unified" = "VCUSD",
    "UNKNOWN"
  )

qualify <- ifelse(
  student_group_count >= 3,
  paste0(
    initials,
    " qualifies as three or more (",
    student_group_count,
    " in this case) have been eligible for DA for at least three out of four consecutive years. "
  ),
  paste(
    initials,
    "does not meet the three or more pupil student group requirement however the following student group(s) have been eligible for DA for at least three out of four consecutive years."
  )
)
```

# `r params$lea` Intensive Assistance

The results of the 2025 California School Dashboard indicate that `r initials` is eligible for Intensive Differentiated Assistance based on the performance of student groups on the Dashboard. Education Code 52072 states that the Solano County Office of Education shall provide intensive assistance when three or more pupil student groups meet the criteria for three or more consecutive years. The California Collaborative for Educational Excellence (CCEE) with the county superintendent of schools shall provide technical assistance to the school district. Using the eligibility criteria, `r initials` qualifies for Intensive Assistance as three or more student groups (`r student_group_count` in this case) have been eligible for DA for at least three out of four consecutive years. Eligible indicators are listed for the previous four years used in qualifying each student group: 2022, 2023, 2024, 2025. The criteria being met by the student group(s) each year may not have been identical from year to year.

```{r da_groups}
#| echo: false

names_groups <- names(clean_groups)

clean_groups |>
  select(-districtname) |>
  flextable() |>
  set_header_labels(values = list("student_group_long" = "Student Group")) |>
  width(width = 1.3) |>
  add_header_row(
    values = c("Student Group", "DA Eligible Indicators"),
    colwidths = c(1, ncol(clean_groups) - 2)
  ) |>
  merge_v(part = "header", j = 1) |>
  theme_vanilla() |>
  align(align = "center", part = "header") |>
  align(align = "left", part = "header", j = 1) |>
  set_caption(paste(
    params$lea,
    "student groups eligible for DA three of the last four years."
  ))


```

```{r student_groups}
#| echo: false
#| warning: false
#| results: asis

wrap_length <- 20

for (group in student_groups_list) {
  academic <-
    dashboard |>
    filter(student_group_long == group) |>
    filter(
      !(indicator %in%
        c(
          "absenteeism",
          "graduation",
          "suspension",
          "college/career",
          "science"
        ))
    ) |>
    ggplot(aes(
      x = reportingyear,
      y = currstatus,
      group = indicator,
      color = indicator
    )) +
    geom_line(linewidth = 1.3) +
    geom_point(size = 2.2) +
    geom_text_repel(
      aes(label = currstatus, color = indicator),
      show.legend = FALSE
    ) +
    scale_x_continuous(breaks = seq(2019, 2025, by = 1)) +
    scale_color_manual(values = indicator_colors) +
    labs(
      title = paste0(
        "Status of **academic indicators** over time for<br>the **",
        group,
        "** student group in ",
        params$lea
      ),
      x = "Year",
      y = "Status (Distance from Standard)",
      color = "Indicator",
      caption = "Colors not reflective of assigned Dashboard colors."
    ) +
    theme_minimal() +
    theme(plot.title = element_markdown(hjust = 0.5, linewidth = wrap_length))

  nonacademic <-
    dashboard |>
    filter(student_group_long == group) |>
    filter(
      indicator %in%
        c("absenteeism", "graduation", "suspension", "college/career")
    ) |>
    ggplot(aes(
      x = reportingyear,
      y = currstatus / 100,
      group = indicator,
      color = indicator
    )) +
    geom_line(linewidth = 1.3) +
    geom_point(size = 2.2) +
    geom_text_repel(
      aes(label = paste0(currstatus, "%"), color = indicator),
      show.legend = FALSE
    ) +
    scale_x_continuous(breaks = seq(2019, 2025, by = 1)) +
    scale_y_continuous(labels = scales::label_percent()) +
    scale_color_manual(values = indicator_colors) +
    labs(
      title = paste0(
        "Status of **non-academic indicators** over time for<br>the **",
        group,
        "** student group in ",
        params$lea
      ),
      x = "Year",
      y = "Status (Percent)",
      color = "Indicator",
      caption = "CCI only reported for 2019, 2023, 2024, and 2025, where applicable."
    ) +
    theme_minimal() +
    theme(plot.title = element_markdown(hjust = 0.5, linewidth = wrap_length))

  print(academic)
  print(nonacademic)
  cat("\n\n")
}

```