---
format: html 
fig-dpi: 300
editor: visual
---

```{r setup, include = FALSE}
library(tidyverse)
library(flextable)
library(openxlsx)

dashboard <- read_csv("data/ca_dashboard.csv") %>%
  filter(
    countyname == "Solano" |
      rtype == "X"
  ) %>%
  mutate(
    statuslevel = as_factor(statuslevel),
    reportingyear = fct_relevel(as_factor(reportingyear), "2019"),
    indicator = as_factor(indicator),
    studentgroup = as_factor(studentgroup),
    student_group_long = fct_relevel(
      as_factor(student_group_long),
      "All students",
      "American Indian or Alaska Native",
      "Asian",
      "Black/African American",
      "Filipino",
      "Hispanic",
      "Pacific Islander",
      "White",
      "Multiple Races/Two or more",
      "Socioeconomically Disadvantaged",
      "Homeless Youth",
      "Students with Disabilities",
      "Foster Youth",
      "English Learner",
      "Long-Term English Learner",
      "English Learners Only",
      "RFEPs Only",
      "English Only",
      "Smarter Balanced Assessment",
      "CA Alternative Assessment"
    ),
    charter_flag = if_else(is.na(charter_flag), "N", charter_flag),
    color = as_factor(case_match(
      color,
      0 ~ "No color",
      1 ~ "Red",
      2 ~ "Orange",
      3 ~ "Yellow",
      4 ~ "Green",
      5 ~ "Blue"
    )),
    lea = if_else(charter_flag == "Y", schoolname, districtname)
  )

status_colors <- c(
  "lightgrey",
  "tomato",
  "orange",
  "yellow",
  "mediumseagreen",
  "dodgerblue",
  "darkgrey"
)

charter_performance_25 <- read.xlsx(
  "data/Charter Performance 2026.xlsx",
  sheet = 1,
  startRow = 4
) |>
  mutate(year = 2025)

charter_performance_24 <- read.xlsx(
  "https://www.cde.ca.gov/sp/ch/documents/perfcat25.xlsx",
  sheet = 1,
  startRow = 4
) |>
  mutate(year = 2024)

charter_performance_23 <- read.xlsx(
  "https://www.cde.ca.gov/sp/ch/documents/perfcat24.xlsx",
  sheet = 1,
  startRow = 4
) |>
  mutate(year = 2023)

charter_performance_19 <- read.xlsx(
  "https://www.cde.ca.gov/sp/ch/documents/ab1505results2020.xlsx",
  sheet = 1,
  startRow = 4
) |>
  mutate(year = 2019)

charter_performance <- bind_rows(
  charter_performance_25,
  charter_performance_24,
  charter_performance_23,
  charter_performance_19
) |>
  mutate(
    Performance_Level = as_factor(Performance_Level)
  ) |>
  mutate(
    Performance_Level = fct_relevel(
      Performance_Level,
      "Low Performing",
      "Middle Performing",
      "High Performing"
    )
  )

charter_info <- read.xlsx(
  "data/chexpirations092625.xlsx",
  sheet = 1,
  startRow = 3
) |>
  janitor::clean_names() |>
  mutate(cds = str_remove_all(county_district_school_code, "-"))
```

## Charter Performance

```{r charter_summary}
#| echo: false

charter_summary <-
  charter_performance %>%
  filter(County == "Solano", year == 2025) %>%
  left_join(
    charter_info,
    by = join_by("CDS" == "cds")
  ) %>%
  select(
    CDS,
    County,
    District,
    School,
    Authorizer = authorizer,
    Performance_Level,
    Criteria,
    `Date Charter Expires` = charter_school_term_expiration_date
  )

```

```{r dashboard_summary}
#| echo: false

exclude_groups <- c(
  "Asian",
  "Filipino",
  "White",
  "Multiple Races/Two or more",
  "Smarter Balanced Assessment"
)
academic_indicators <- c("ELA", "Math", "ELPI", "college/career")

dashboard_summary <-
  dashboard %>%
  filter(
    reportingyear %in% c("2025", "2024", "2023"),
    (countyname == "Solano" & charter_flag == "Y") |
      (rtype == "X"),
    indicator %in% academic_indicators,
    !(student_group_long %in% exclude_groups),
    color != "No color"
  ) %>%
  select(reportingyear, schoolname, indicator, student_group_long, currstatus)

ca_status <-
  dashboard_summary %>%
  filter(schoolname == "State of California") %>%
  rename(ca_status = currstatus) %>%
  select(-schoolname)

solano_status <-
  dashboard_summary %>%
  filter(schoolname != "State of California") %>%
  left_join(
    ca_status,
    by = join_by(reportingyear, indicator, student_group_long)
  )

solano_groups_total <-
  solano_status %>%
  group_by(schoolname, indicator) %>%
  summarise(total_groups = n())

solano_groups_ca_compare_count <-
  solano_status %>%
  mutate(
    compare_ca = case_when(
      currstatus < ca_status ~ "Below",
      currstatus == ca_status ~ "Equal",
      currstatus > ca_status ~ "Above"
    )
  ) %>%
  group_by(schoolname, indicator, compare_ca) %>%
  count() %>%
  left_join(solano_groups_total, by = join_by(schoolname, indicator))

solano_indicators <-
  charter_summary %>%
  select(School, Performance_Level, Criteria) %>%
  left_join(solano_status, by = join_by(School == schoolname)) %>%
  left_join(
    solano_groups_ca_compare_count,
    by = join_by(School == schoolname, indicator)
  ) %>%
  select(
    School,
    Performance_Level,
    Criteria,
    indicator,
    total_groups,
    compare_ca,
    n,
    reportingyear,
    student_group_long,
    ca_status,
    currstatus
  ) %>%
  filter(
    (compare_ca == "Below" & currstatus < ca_status) |
      (compare_ca == "Equal" & currstatus == ca_status) |
      (compare_ca == "Above" & currstatus > ca_status)
  ) %>%
  arrange(School, indicator, compare_ca, reportingyear, student_group_long)

# Doesn't work exactly how I want it, but it's close
# solano_indicators %>%
#   flextable() %>%
#   merge_v(j = 1:3, combine = T) %>%
#   merge_v(j = 4:5, combine = T) %>%
#   merge_v(j = 6:7, combine = T) %>%
#   merge_v(j = 8, combine = T) %>%
#   theme_vanilla()

# solano_status %>%
#   ggplot() +
#   geom_hline(aes(yintercept = ca_status), color = "#2B419B") +
#   geom_point(aes(x = student_group_long, y = currstatus,
#                  color = schoolname), show.legend = F) +
#   facet_grid(schoolname ~ indicator, switch = "y") +  # Moves strip labels to the left
#   labs(title = "Charter Student Groups in relation to CA Status for Academic indicators",
#        x = NULL,
#        y = "Status") +
#   theme_minimal() +
#   theme(axis.text.x = element_blank(),
#         axis.text.y = element_blank(),
#         strip.text.y.left = element_text(angle = 0))

```

```{r write_excel}

# Filter data
filtered_charter <-
  charter_performance %>%
  filter(County == "Solano")

filtered_dashboard <-
  dashboard %>%
  filter(
    reportingyear %in% c("2024", "2023"),
    (countyname == "Solano" & charter_flag == "Y") |
      (rtype == "X")
  )

# Add row indices to the dataset before any grouping operations
solano_indicators <- solano_indicators %>%
  mutate(row_idx = row_number())

# Get merge ranges for hierarchical levels
get_merge_ranges <- function(data, group_vars) {
  data %>%
    group_by(across(all_of(group_vars))) %>%
    summarise(
      start_row = min(row_idx) + 1, # +1 for header row
      end_row = max(row_idx) + 1,
      .groups = "drop"
    ) %>%
    filter(start_row < end_row) # Only include groups that need merging (more than 1 row)
}

# Calculate all merge ranges at different hierarchy levels
merge_school <- get_merge_ranges(solano_indicators, c("School"))
merge_perf <- get_merge_ranges(
  solano_indicators,
  c("School", "Performance_Level")
)
merge_criteria <- get_merge_ranges(
  solano_indicators,
  c("School", "Performance_Level", "Criteria")
)
merge_indicator <- get_merge_ranges(
  solano_indicators,
  c("School", "Performance_Level", "Criteria", "indicator")
)
merge_groups <- get_merge_ranges(
  solano_indicators,
  c("School", "Performance_Level", "Criteria", "indicator", "total_groups")
)
merge_compare <- get_merge_ranges(
  solano_indicators,
  c(
    "School",
    "Performance_Level",
    "Criteria",
    "indicator",
    "total_groups",
    "compare_ca"
  )
)
merge_n <- get_merge_ranges(
  solano_indicators,
  c(
    "School",
    "Performance_Level",
    "Criteria",
    "indicator",
    "total_groups",
    "compare_ca",
    "n"
  )
)
merge_year <- get_merge_ranges(
  solano_indicators,
  c(
    "School",
    "Performance_Level",
    "Criteria",
    "indicator",
    "total_groups",
    "compare_ca",
    "n",
    "reportingyear"
  )
)

# Create Excel Workbook
wb <- createWorkbook()
addWorksheet(wb, "Charter Summary")
addWorksheet(wb, "Dashboard summary (Merged)")
addWorksheet(wb, "Dashboard summary (Unmerged)")
addWorksheet(wb, "Charter Performance ORIGINAL")
addWorksheet(wb, "CA Dashboard (Charters)")

# Add custom style for merged cells
headerStyle <- createStyle(
  fontColour = "#FFFFFF",
  fgFill = "#4F81BD",
  halign = "center",
  valign = "center",
  textDecoration = "bold",
  border = "TopBottomLeftRight"
)

# Write data - remove row_idx column before writing
writeData(wb, "Charter Summary", charter_summary)
writeData(wb, "Dashboard summary (Merged)", select(solano_indicators, -row_idx))
writeData(
  wb,
  "Dashboard summary (Unmerged)",
  select(solano_indicators, -row_idx)
)
writeData(wb, "Charter Performance ORIGINAL", filtered_charter)
writeData(wb, "CA Dashboard (Charters)", filtered_dashboard)

# Apply styles to header row
addStyle(
  wb,
  "Dashboard summary (Merged)",
  headerStyle,
  rows = 1,
  cols = 1:(ncol(solano_indicators) - 1)
)

# Function to apply merges for a specific column
apply_column_merges <- function(wb, sheet, merge_data, column_name, data) {
  col_index <- which(names(data) == column_name)
  if (length(col_index) == 1 && nrow(merge_data) > 0) {
    for (i in 1:nrow(merge_data)) {
      mergeCells(
        wb,
        sheet,
        cols = col_index,
        rows = merge_data$start_row[i]:merge_data$end_row[i]
      )
    }
  }
}

# Apply merges for each column based on their respective hierarchical level
apply_column_merges(
  wb,
  "Dashboard summary (Merged)",
  merge_school,
  "School",
  solano_indicators
)
apply_column_merges(
  wb,
  "Dashboard summary (Merged)",
  merge_perf,
  "Performance_Level",
  solano_indicators
)
apply_column_merges(
  wb,
  "Dashboard summary (Merged)",
  merge_criteria,
  "Criteria",
  solano_indicators
)
apply_column_merges(
  wb,
  "Dashboard summary (Merged)",
  merge_indicator,
  "indicator",
  solano_indicators
)
apply_column_merges(
  wb,
  "Dashboard summary (Merged)",
  merge_groups,
  "total_groups",
  solano_indicators
)
apply_column_merges(
  wb,
  "Dashboard summary (Merged)",
  merge_compare,
  "compare_ca",
  solano_indicators
)
apply_column_merges(
  wb,
  "Dashboard summary (Merged)",
  merge_n,
  "n",
  solano_indicators
)
apply_column_merges(
  wb,
  "Dashboard summary (Merged)",
  merge_year,
  "reportingyear",
  solano_indicators
)

# Auto-size columns for better readability
setColWidths(
  wb,
  "Dashboard summary (Merged)",
  cols = 1:(ncol(solano_indicators) - 1),
  widths = "auto"
)

# Add formatting for easier reading
addFilter(
  wb,
  "Dashboard summary (Merged)",
  row = 1,
  cols = 1:(ncol(solano_indicators) - 1)
)
freezePane(wb, "Dashboard summary (Merged)", firstRow = TRUE)

# Save workbook
saveWorkbook(wb, "Solano charter performance 2026.xlsx", overwrite = TRUE)
```

## Charter Performance Levels Over Time

The following chart displays the CDE performance levels for `r school_printing` from 2023 to 2025. These levels are determined by the CA Department of Education based on student outcomes on CA State Dashboard indicators. Charter schools were not assigned performance levels in 2020, 2021, or 2022.

```{r lea_chart}
#| echo: false

library(ggrepel)

charter_list <- c(
  "Caliber: ChangeMakers Academy",
  "Dixon Montessori Charter",
  "Elite Public",
  "Griffin Academy High",
  "Kairos Public",
  "MIT Academy",
  "MIT Griffin Academy Middle",
  "Mare Island Technology Academy"
)

charter_performance |>
  filter(County == "Solano", School %in% charter_list, year >= 2023) |>
  ggplot(aes(x = year, y = Performance_Level)) +
  geom_label_repel(aes(label = School)) +
  scale_y_discrete(drop = FALSE) +
  scale_x_continuous(breaks = c(2023, 2024, 2025)) +
  labs(
    title = paste("Performance Levels Over Time for", school_printing),
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 16, hjust = 0.5)
  )

```

```{r}
#| label: ai-version
#| echo: false

# Ensure Performance_Level is an ordered factor to maintain the 3x3 grid structure
# Adjust levels to match your specific CDE data labels (e.g., Blue, Green, Yellow)
charter_performance_processed <- charter_performance |>
  filter(County == "Solano", School %in% charter_list, year >= 2023) |>
  mutate(
    Performance_Level = fct_relevel(
      Performance_Level,
      "High Performing",
      "Middle Performing",
      "Low Performing",
    )
  ) |>
  # mutate(School = str_wrap(School, width = 20)) |>
  group_by(year, Performance_Level) |>
  mutate(v_offset = row_number()) |>
  ungroup()

# Define the color mapping based on school branding/logos
charter_colors <- c(
  "Caliber: ChangeMakers Academy" = "#043689", # Navy
  "Dixon Montessori Charter" = "#d0ecef", # Forest Green
  "Elite Public" = "#9966ff", # Golden Yellow
  "Griffin Academy High" = "#6db5b9", # Teal
  "Kairos Public" = "#012c60", # Midnight Blue
  "MIT Academy" = "#0f0f38", # Navy Blue
  "MIT Griffin Academy Middle" = "#6db5b9", # Teal (Griffin Brand)
  "Mare Island Technology Academy" = "#0f0f38" # Navy Blue (MIT Brand)
)

# Define the color mapping based on school branding/logos
charter_text_colors <- c(
  "Caliber: ChangeMakers Academy" = "white", # Navy
  "Dixon Montessori Charter" = "black", # Forest Green
  "Elite Public" = "black", # Golden Yellow
  "Griffin Academy High" = "black", # Teal
  "Kairos Public" = "white", # Midnight Blue
  "MIT Academy" = "white", # Navy Blue
  "MIT Griffin Academy Middle" = "black", # Teal (Griffin Brand)
  "Mare Island Technology Academy" = "white" # Navy Blue (MIT Brand)
)

ggplot(charter_performance_processed, aes(x = 1, y = v_offset)) +
  geom_label(
    aes(label = School, fill = School, color = School),
    size = 6.5,

    show.legend = FALSE
  ) +
  facet_grid(Performance_Level ~ year, switch = "y") +
  scale_x_continuous(expand = expansion(mult = 0.5)) +
  scale_y_continuous(expand = expansion(add = 0.75)) +
  scale_fill_manual(values = charter_colors) +
  scale_color_manual(values = charter_text_colors) +
  labs(
    title = "Solano County Charter Performance (2023-2025)",
    x = "Year",
    y = "CDE Performance Level"
  ) +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA),
    strip.background = element_rect(fill = "gold"),
    axis.title = element_text(size = 16),
    strip.text = element_text(size = 14),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
  )

```


## CA State Dashboard Results

```{r lea_table}
#| echo: false

indicator_summaries <-
  dashboard |>
  select(
    reportingyear,
    countyname,
    districtname,
    schoolname,
    student_group_long,
    indicator,
    currstatus,
    change,
    color
  ) |>
  mutate(
    narative = case_match(
      indicator,
      "ELA" ~ paste0(
        abs(currstatus),
        if_else(currstatus < 0, " points below", " points above"),
        " standard in ELA. (",
        if_else(change < 0, "decreased ", "increased "),
        abs(change),
        " points)"
      ),
      "Math" ~ paste0(
        abs(currstatus),
        if_else(currstatus < 0, " points below", " points above"),
        " standard in Math. (",
        if_else(change < 0, "decreased ", "increased "),
        abs(change),
        " points)"
      ),
      "absenteeism" ~ paste0(
        currstatus,
        "% chronically absent (",
        if_else(change < 0, "decreased ", "increased "),
        abs(change),
        "%)"
      ),
      "suspension" ~ paste0(
        currstatus,
        "% suspended at least one day (",
        if_else(change < 0, "decreased ", "increased "),
        abs(change),
        "%)"
      ),
      "graduation" ~ paste0(
        currstatus,
        "% graduated (",
        if_else(change < 0, "decreased ", "increased "),
        abs(change),
        "%)"
      ),
      "college/career" ~ paste0(
        currstatus,
        "% prepared (",
        if_else(change < 0, "decreased ", "increased "),
        abs(change),
        "%)"
      ),
      "science" ~ paste0(
        currstatus,
        " science points (",
        if_else(change < 0, "decreased ", "increased "),
        abs(change),
        " points)"
      ),
      .default = "No data available"
    ),
    color_hex = case_match(
      color,
      "Red" ~ "#ffa0a0ff",
      "Orange" ~ "#fdd5a1ff",
      "Yellow" ~ "#ffff9fff",
      "Green" ~ "#aae4a8ff",
      "Blue" ~ "#99bdffff",
      "No color" ~ "#ccccccff",
      .default = "#FFFFFF"
    ),
    indicator = case_match(
      indicator,
      "ELA" ~ "English Language Arts (ELA)",
      "Math" ~ "Mathematics",
      "absenteeism" ~ "Chronic Absenteeism",
      "suspension" ~ "Suspension Rate",
      "graduation" ~ "Graduation Rate",
      "college/career" ~ "College/Career Readiness (CCI)",
      "science" ~ "Science"
    )
  )

```

### California Results for All Students

```{r}
#| label: CA_2025_dashboard_table
#| echo: false

ca_summaries <-
  indicator_summaries |>
  filter(
    schoolname == "State of California",
    reportingyear == "2025",
    student_group_long == "All students"
  ) |>
  select(indicator, narative, color_hex)

ca_summaries |>
  flextable() |>
  bg(j = "narative", bg = ca_summaries$color_hex, part = "body") |>
  delete_columns(j = "color_hex") |>
  set_header_labels(
    indicator = "CA State Dashboard Indicator",
    narative = "State of CA Public Dashboard Results"
  ) |>
  width(width = c(2.5, 4.5)) |>
  theme_vanilla()

```

The following table shows the CA State Dashboard results for all students at `r school_printing` for the past three years. The background shading of each cell represents the color 

### `r school_printing` Results for All Students

```{r}
#| label: 2025_dashboard_table
#| echo: false

# 1. Prepare pivoted data
charter_summaries <- indicator_summaries |>
  filter(
    schoolname == params$school,
    reportingyear %in% c("2023", "2024", "2025"),
    student_group_long == "All students"
  ) |>
  select(reportingyear, indicator, currstatus, color_hex) |>
  pivot_wider(
    names_from = reportingyear,
    values_from = c(currstatus, color_hex),
    names_sep = "_"
  )

# 2. Build flextable with targeted coloring
charter_summaries |>
  select(
    indicator,
    currstatus_2023,
    currstatus_2024,
    currstatus_2025,
    color_hex_2023,
    color_hex_2024,
    color_hex_2025
  ) |>
  flextable() |>
  theme_vanilla() |>
  # Color the 2023 Status column using the 2023 Color column
  bg(
    j = "currstatus_2023",
    bg = charter_summaries$`color_hex_2023`,
    part = "body"
  ) |>
  # Color the 2024 Status column using the 2024 Color column
  bg(
    j = "currstatus_2024",
    bg = charter_summaries$`color_hex_2024`,
    part = "body"
  ) |>
  # Color the 2025 Status column using the 2025 Color column
  bg(
    j = "currstatus_2025",
    bg = charter_summaries$`color_hex_2025`,
    part = "body"
  ) |>
  # Remove both color helper columns
  delete_columns(j = c("color_hex_2023", "color_hex_2024", "color_hex_2025")) |>
  set_header_labels(
    indicator = "CA State Dashboard Indicator",
    `currstatus_2023` = "2023 Status",
    `currstatus_2024` = "2024 Status",
    `currstatus_2025` = "2025 Status"
  ) |>
  width(width = c(2.5, 1, 1, 1)) |>
  theme_vanilla()

```