---
format:
  docx:
    fig-dpi: 300
    fig-width: 7.5
    reference-doc: scoe-are-report.docx
editor: visual
params:
  lea: Vacaville Unified
---

```{r setup}
#| include: false

library(tidyverse)
library(flextable)
library(ggrepel)
library(ggtext)
library(here)

if (interactive()) {
  params <- rmarkdown::yaml_front_matter(
    rstudioapi::getSourceEditorContext()$path
  )$params
}

dashboard <- read_csv(here("data/solano_dashboard.csv")) |>
  filter(
    rtype == "D" &
      districtname == params$lea |
      charter_flag == "Y" & rtype == "S" & schoolname == params$lea
  ) |>
  select(
    reportingyear,
    districtname,
    schoolname,
    student_group_long,
    indicator,
    currdenom,
    currstatus,
    color,
    change,
    indicator_eligible
  )

indicator_colors = c(
  "ELA" = "#CC6677",
  "Math" = "#882255",
  "ELPI" = "#AA4499",
  "college/career" = "#332288",
  "graduation" = "#88CCEE",
  "absenteeism" = "#44AA99",
  "suspension" = "#117733",
  "science" = "#DDCC77"
)

da_groups <-
  dashboard |>
  mutate(reportingyear = as_factor(reportingyear)) |>
  filter(
    reportingyear %in% c(2025, 2024, 2023),
    indicator_eligible == TRUE
  ) |>
  select(-indicator_eligible)

long_da_groups <-
  da_groups |>
  select(
    reportingyear,
    districtname,
    schoolname,
    student_group_long,
    indicator
  ) |>
  arrange(reportingyear) |>
  pivot_wider(
    names_from = reportingyear,
    values_from = indicator,
    values_fn = list
  ) |>
  arrange(districtname, schoolname, student_group_long) |>
  mutate(
    across(where(is.list), ~ map_chr(., ~ paste(., collapse = "; "))),
    da_years = rowSums(across(matches("^\\d{4}$"), ~ . != ""))
  ) |>
  filter(da_years >= 3) |>
  select(districtname, student_group_long, `2023`, `2024`, `2025`)

```

```{r create_variables}
#| include: false

student_groups_list <- unique(long_da_groups$student_group_long)
student_group_count <- length(student_groups_list)

initials <-
  switch(
    params$lea,
    "Solano County Office of Education" = "SCOE",
    "Benicia Unified" = "BUSD",
    "Dixon Unified" = "DUSD",
    "Fairfield-Suisun Unified" = "FSUSD",
    "Travis Unified" = "TUSD",
    "Vacaville Unified" = "VUSD",
    "Vallejo City Unified" = "VCUSD",
    "UNKNOWN"
  )
```

# `r params$lea` Targeted Assistance

The results of the 2025 California School Dashboard indicate `r initials` is eligible for Targeted Differentiated Assistance based on the performance of student groups on the Dashboard. Education Code 52071(f) states that the Solano County Office of Education shall provide targeted assistance when one or more pupil student groups meet the criteria pursuant for three or more consecutive years. The geographic lead agency shall collaborate with the county superintendent of schools and provide technical assistance to the school district. Using the eligibility criteria, `r initials` has qualified for Targeted Assistance based on the same student groups who meet the criteria for three or more consecutive years:

```{r da_groups}
#| echo: false

names_groups <- names(long_da_groups)

long_da_groups |>
  select(-districtname) |>
  flextable() |>
  set_header_labels(values = list("student_group_long" = "Student Group")) |>
  width(width = 1.3) |>
  add_header_row(
    values = c("Student Group", "DA Eligible Indicators"),
    colwidths = c(1, ncol(long_da_groups) - 2)
  ) |>
  merge_v(part = "header", j = 1) |>
  theme_vanilla() |>
  align(align = "center", part = "header") |>
  align(align = "left", part = "header", j = 1) |>
  set_caption(paste(
    params$lea,
    "student groups eligible for DA the last three years."
  ))


```

```{r student_groups}
#| echo: false
#| warning: false
#| results: asis

wrap_length <- 20

for (group in student_groups_list) {
  academic <-
    dashboard |>
    filter(student_group_long == group) |>
    filter(indicator %in% c("ELA", "Math")) |>
    drop_na()

  min_year <- min(academic$reportingyear)
  max_year <- max(academic$reportingyear)

  academic_graph <-
    academic |>
    ggplot(aes(
      x = reportingyear,
      y = currstatus,
      group = indicator,
      color = indicator
    )) +
    geom_line(linewidth = 1.3) +
    geom_point(size = 2.2) +
    geom_text_repel(
      aes(label = currstatus, color = indicator),
      show.legend = FALSE
    ) +
    scale_color_manual(values = indicator_colors) +
    scale_x_continuous(breaks = seq(min_year, max_year, by = 1)) +
    labs(
      title = paste0(
        "Status of **academic indicators** over time for<br>the **",
        group,
        "** student group in ",
        params$lea
      ),
      x = "Year",
      y = "Status (Distance from Standard)",
      color = "Indicator",
      caption = "Colors not reflective of assigned Dashboard colors."
    ) +
    theme_minimal() +
    theme(plot.title = element_markdown(hjust = 0.5, linewidth = wrap_length))

  print(academic_graph)

  if (group %in% c("English Learner", "Long-Term English Learner")) {
    elpi <-
      dashboard |>
      filter(student_group_long == group) |>
      filter(indicator == "ELPI")

    min_year <- min(elpi$reportingyear)
    max_year <- max(elpi$reportingyear)

    elpi_graph <-
      elpi |>
      ggplot(aes(
        x = reportingyear,
        y = currstatus,
        group = indicator,
        color = indicator
      )) +
      geom_line(linewidth = 1.3) +
      geom_point(size = 2.2) +
      geom_text_repel(
        aes(label = currstatus, color = indicator),
        show.legend = FALSE
      ) +
      scale_color_manual(values = indicator_colors) +
      scale_x_continuous(breaks = seq(min_year, max_year, by = 1)) +
      labs(
        title = paste0(
          "Status of ELPI over time for<br>the **",
          group,
          "** student group in ",
          params$lea
        ),
        x = "Year",
        y = "Status (Percent making progress)",
        color = "Indicator",
        caption = "Colors not reflective of assigned Dashboard colors."
      ) +
      theme_minimal() +
      theme(plot.title = element_markdown(hjust = 0.5, linewidth = wrap_length))

    print(elpi_graph)
  }

  nonacademic <-
    dashboard |>
    filter(student_group_long == group) |>
    filter(
      indicator %in%
        c("absenteeism", "graduation", "suspension", "college/career")
    )

  min_year <- min(nonacademic$reportingyear)
  max_year <- max(nonacademic$reportingyear)

  nonacademic_graph <-
    nonacademic |>
    ggplot(aes(
      x = reportingyear,
      y = currstatus / 100,
      group = indicator,
      color = indicator
    )) +
    geom_line(linewidth = 1.3) +
    geom_point(size = 2.2) +
    geom_text_repel(
      aes(label = paste0(currstatus, "%"), color = indicator),
      show.legend = FALSE
    ) +
    scale_y_continuous(labels = scales::label_percent()) +
    scale_x_continuous(breaks = seq(min_year, max_year, by = 1)) +
    scale_color_manual(values = indicator_colors) +
    labs(
      title = paste0(
        "Status of **non-academic indicators** over time for<br>the **",
        group,
        "** student group in ",
        params$lea
      ),
      x = "Year",
      y = "Status (Percent)",
      color = "Indicator",
      caption = "CCI only reported for 2019, 2023, 2024, and 2025, where applicable."
    ) +
    theme_minimal() +
    theme(plot.title = element_markdown(hjust = 0.5, linewidth = wrap_length))

  print(nonacademic_graph)
  cat("\n\n")
}

```